---
title: "Project_One"
author: "John K. Hancock"
date: "September 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(warn=-1)
library(XML)
library(httr)
library(curl)
library(RCurl)
library(stringr)
library(tm)
library(sm)
```

####DATA607 Fall 2018  - PROJECT ONE
####John K. Hancock

1. Read in the text file, collapse it as one large block of text, unlist and parse it into a vector using str_split, and go through a series of replace statements to search for strings and replace them.


```{r}
txt<- character()
txt <- c(txt, readLines('tournamentinfo.txt'))
txt <- str_c(txt, collapse = "\n")
txt_parsed <-unlist(str_split(txt, "-----------------------------------------------------------------------------------------"))
txt_parsed<-str_replace(txt_parsed,pattern="^\n","")
txt_parsed<-str_replace(txt_parsed,pattern="\n$","")
txt_parsed<-str_replace(txt_parsed," \n Num  ","Num")
txt_parsed<-str_replace(txt_parsed," / R:","|")
txt_parsed<-str_replace(txt_parsed,"USCF ID / Rtg ", "USCF ID|Rtg")
txt_parsed<-str_replace(txt_parsed,"->","|")
txt_parsed<-str_replace(txt_parsed,"Rtg\\(Pre|Post\\)", "Rating Prior|Rating Post")
txt_parsed<-str_replace(txt_parsed,"Post\\)", "")
txt_parsed<-str_replace(txt_parsed,"\n","")
txt_parsed<-txt_parsed[txt_parsed != ""]


```

2. The first line below shows the second element in the txt_parsed vector.  The next statement locates the coordinates for  where all of the "|" delimiters are located. 

```{r}
txt_parsed[2]
str_locate_all(txt_parsed[2], "\\|")
```

3. With the coordinates of the "|", the next step is to extract the data into individual vectors, player_id, player_name, player_state, total_no_of_points, and player_pre_rating.  The vector, player_opp, contains the player_ids for a player's opponent.  The vector black_white contains each instance of whether the player's chess pieces were Black or White. 



```{r}
player_id <- trimws(str_sub(txt_parsed[2:65],0,6))
player_name<- trimws(str_sub(txt_parsed[2:65],8,40))
player_state<- trimws(str_sub(txt_parsed[2:65],90,95))
total_no_of_points<-trimws(str_sub(txt_parsed[2:65],42,46))
player_pre_rating<-str_replace(trimws(str_sub(txt_parsed[2:65],107,114)),"P.*$","") 
player_opp<-trimws(str_sub(txt_parsed[2:65],48,88))
black_white <- trimws(str_sub(txt_parsed[2:65],132,168))
```

```{r}
player_id<-as.numeric(player_id)
player_id
```





```{r}
player_name
```

```{r}
player_state
```

```{r}
total_no_of_points<-as.numeric(total_no_of_points)
total_no_of_points
```

```{r}
player_pre_rating<-as.numeric(player_pre_rating)
player_pre_rating
```

4. The vectors player_id, player_name, player_state, total_no_of_points, and player_pre_rating were combined into a data frame. 



```{r}

df = data.frame(player_id,player_name,player_state,total_no_of_points,player_pre_rating)
df
```

5. In the next block, I created a test to calculate the Average Pre Chess Rating of Opponents. This is done by extracting the player_ids of opponents by searching for only 1 to 2 digits in the string.  Next these were added to a numeric vector.  With the player_ids, I created a temporary data frame for a player's opponent by subsetting records for only the opponents.  Finally, by taking the mean of these players, I have each the Average Pre Chess Rating of Opponents. 


```{r}
test<- unlist(str_extract_all(player_opp[1],"[:digit:]{1,2}"))
test <- as.numeric(test)
temp<- df[df$player_id %in% test,]
mean(temp$player_pre_rating)

```

5a. With a successful test, I created a for loop which calculates the Average Pre Chess Rating of Opponents for all of the playerrs, and then adding that vector to the data frame. 



```{r}
opponent_avg <- vector(mode="numeric")

for (i in seq(1,length(player_opp)))
{
	
	test<- unlist(str_extract_all(player_opp[i],"[:digit:]{1,2}"))
	test <- as.numeric(test)
	temp<- df[df$player_id %in% test,]
	opponent_avg<-append(opponent_avg, mean(temp$player_pre_rating))
}

opponent_avg




```

```{r}
df$avg_pre_chess_rating_of_opponents<-c(opponent_avg)
```


```{r}
df
```

6. I went a step further.  I wanted to know how many Wins, Losses, Draws, U's, and H's that a player has.  Additionally, I wanted to count the number of times a player has played with Black or White chess pieces. Using extract all on player_opp and black_white to search for "W","L","D","U", and "H" and putting these values are put into vectors, I was able to count the length of each vector to calculate the number of Wins, Losses, Draws, Us, and Hs.  I took the same approach to count the number of times a player played with either Black or White pieces. 



```{r}
countofWins<- vector(mode="numeric")
countofLosses<-vector(mode="numeric")
countofDraws <-vector(mode="numeric")
countofU <-vector(mode="numeric")
countofH <-vector(mode="numeric")
countofWhite <-vector(mode="numeric")
countofBlack <- vector(mode="numeric")

for (i in seq(1,length(player_opp)))
{
		
	wins<- unlist(str_extract_all(player_opp[i],"[W]"))
	countofWins<-append(countofWins,length(wins))
	
	losses <- unlist(str_extract_all(player_opp[i],"[L]"))
	countofLosses<-append(countofLosses,length(losses))
	
	draws <- unlist(str_extract_all(player_opp[i],"[D]"))
	countofDraws<-append(countofDraws,length(draws))
	
	U <- unlist(str_extract_all(player_opp[i],"[U]"))
	countofU<-append(countofU,length(U))
	
	H<-unlist(str_extract_all(player_opp[i],"[H]"))
	countofH<-append(countofH,length(H))
	
	WHite<-unlist(str_extract_all(black_white[i],"[W]"))
	countofWhite<-append(countofWhite,length(WHite))
	
	Black<-unlist(str_extract_all(black_white[i],"[B]"))
	countofBlack<-append(countofBlack,length(Black))
}


```

7. These values were added to the data frame. 


```{r}
df$Wins <- countofWins
df$Losses <- countofLosses
df$Draws <- countofDraws 
df$U <- countofU 
df$H <- countofH
df$White<- countofWhite
df$Black<-countofBlack
df

```


```{r}
summary(df)
```
8. Finally, write the results to a csv file for loading into a database.


```{r}
write.csv(df, file = "Chess Tournament Results.csv", row.names = FALSE)
```






