---
title: "DATA607 WEEK One Assignment"
author: "John K. Hancock"
date: "September 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### DATA606 WEEK ONE ASSIGMENT


```{r}
options(warn=-1)
setwd('C:/Users/jkhan/Documents/CUNY/Fall 2018/DATA606/Week One due 09-02-2018')
library(XML)
library(httr)
library(curl)
library(stringr)
```
### Use Curl to download the data to a csv file.  Read the csv file into a dataframe.

```{r}
curl_download('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data', 'mushrooms.csv')
mushrooms <- read.csv('mushrooms.csv',header = FALSE)
unlink('mushrooms.csv')
```
```{r}
head(mushrooms,5)
```

##Download the dictionary defining the columns.


```{r}
dictionary <- curl_download('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names', 'dictionary.txt')
dictionary_text <- readLines('dictionary.txt')

```

### Use grep to find the Attribute section of the page
```{r}
grep('Attribute', dictionary_text)
```
```{r}
vec <- c(dictionary_text[107:139])
vec 
```
### Use a regular expression to find everything between the period and the colon.  Remove extraneous values.

```{r}
vec2 = gsub(".*\\.(.*)\\:.*", "\\1",vec)
vec2 <-vec2[-c(2,5,8,13,14,17,21,23,28,30,32)]

```

### Trim the whitespaces 

```{r}
vec3 <- trimws(vec2)
vec4 <- c("edible_poisonous")
column_names <- c(vec4,vec3)
length(column_names)
```
### Print out the first 7 rows of the data frame and take a look at its structure and names



```{r}
colnames(mushrooms) <- c(column_names)

head(mushrooms,7)
```
```{r}
str(mushrooms)
```

names
```{r}
names(mushrooms)
```

### Next I subsetted a new data frame using columns, "edible_poisonous", "odor", "population", and "habitat"



```{r}
new_shrooms <- mushrooms[c(1,6,22:23)]
str(new_shrooms)
```

```{r}
names(new_shrooms)
```

### To convert the abbreviations, I had to change the column to a character, convert the abbreviations, and then re-convert the column back to factor.

```{r}
new_shrooms$edible_poisonous <- as.character(new_shrooms$edible_poisonous)
new_shrooms$edible_poisonous[new_shrooms$edible_poisonous == "p"] <- "poisonous"
new_shrooms$edible_poisonous[new_shrooms$edible_poisonous == "e"] <- "edible"
new_shrooms$edible_poisonous <- as.factor(new_shrooms$edible_poisonous)


```
```{r}
str(new_shrooms)
```

### Repeat the process for the odor observation, but use a switch statement

```{r}
head(new_shrooms$odor,10)
```

```{r}
new_shrooms$odor<- as.character(new_shrooms$odor)
new_shrooms$odor <- sapply(new_shrooms$odor, switch,
                   	a='almond',
					l='anise',
 					c='creosote',
 					y='fishy',
 					f='foul',
 					m='musty',
 					n='none',
 					p='pungent',
 					s='spicy')
new_shrooms$odor<- as.factor(new_shrooms$odor)

```
```{r}
head(new_shrooms$odor,10)
```

```{r}
str(new_shrooms)
```
```{r}
new_shrooms$population<- as.character(new_shrooms$population)
new_shrooms$population <- sapply(new_shrooms$population, switch,
                   	a='abundant',
					c='clustered',
 					n='numerous',
 					s='scattered',
					v='several',
					y='solitary')
new_shrooms$population<- as.factor(new_shrooms$population)

```
```{r}
head(new_shrooms$population,10)
```
### Finally, the habitat observation
```{r}
new_shrooms$habitat<- as.character(new_shrooms$habitat)
new_shrooms$habitat <- sapply(new_shrooms$habitat, switch,
                   	 g='grasses',
					 l='leaves',
					 m='meadows',
					 p='paths',
				 	 u='urban',
					 w='waste',
					 d='woods')
new_shrooms$habitat<- as.factor(new_shrooms$habitat)
```
```{r}
head(new_shrooms$habitat,10)
```
```{r}
str(new_shrooms)
```

```{r}
head(new_shrooms,20)
```

