---
title: "Presentation"
author: "John K. Hancock"
date: "September 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###Presentation R and RDBMS

####Using the RSQLite and DBI packages
```{r}
options(warn=-1)
library(RSQLite)
library(DBI)
library(plyr)
library(ggplot2)
library(reshape2)


```
####Connecting  to the a SQLite version of the Lahman Baseball Database
```{r}
con = dbConnect(RSQLite::SQLite(),'lahman2016.sqlite')
```

####List Tables
```{r}
dbListTables(con)
```
####Master Table contains bibliographic information about the players

```{r}
dbListFields(con, "Master")
```
####Pitching Table contains information about pitching data.  These tables are related by playerID

```{r}
dbListFields(con, "Pitching")
```
####Select Pitchers with 3,000 or more strikeouts by using a SQL statement to join the Master and Pitching tables 

```{r}
pitchers<- dbGetQuery(con, "SELECT Master.nameLast, Master.nameGiven, sum(Pitching.SO) AS 'Strikeouts', sum(Pitching.W) As 'Wins'
                                            FROM Master  
                                            INNER JOIN Pitching ON Pitching.playerID = Master.playerID
                                            GROUP BY Master.nameLast, Master.nameGiven 
                                            HAVING sum(Pitching.SO) >= 3000
                                            Order BY sum(Pitching.SO) desc, sum(Pitching.W) desc ")
                                            
```

```{r}
pitchers
```
```{r}
top_3_pitchers<- dbGetQuery(con, "SELECT Master.nameLast || ', ' || Master.nameGiven AS 'Name', sum(Pitching.SO) AS 'Strikeouts',     sum(Pitching.W) As                                  'Wins'
                                            FROM Master  
                                            INNER JOIN Pitching ON Pitching.playerID = Master.playerID
                                            GROUP BY Master.nameLast, Master.nameGiven 
                                            HAVING sum(Pitching.SO) >= 3000
                                            Order BY sum(Pitching.SO) desc, sum(Pitching.W) desc 
                                            Limit 3")
```

```{r}
top_3_pitchers
```

```{r}
data.m <- melt(top_3_pitchers, id.vars='Name')

# plot everything
ggplot(data.m, aes(Name, value)) +   
  geom_bar(aes(fill = variable), position = "dodge", stat="identity")+
geom_text(aes(label=value), vjust=1.6, hjust=-.25, color="blue",
            position = position_dodge(1.0), size=3.5)
  


```

